---
title: "Homework 1 BACS"
author: "109006206"
date: "2023-02-19"
output: pdf_document
fontsize: 12pt
header-includes:
  - \usepackage{setspace}
  - \onehalfspacing
  - \usepackage{fancyhdr}
  - \fancyhead[r]{109006206}
  - \pagestyle{fancy}
  - \setlength{\headheight}{15pt}
---

### Opening Libraries
```{r,message=FALSE}
require(dplyr)
```
<br>

### Set Working Directories
```{r,message=FALSE}
setwd("/Users/olivia/Documents/Documents/Study/Semester 6/BACS")
```
<br>

### Import the Data
```{r,message=FALSE}
data<-read.table(file="customers.txt",header=TRUE)
```
<br>

#### 1. What is the 5th element in the original list of ages?
##### To get the 5th element we need to access the 5th row of the data so here i use data[row,column]
```{r,message=FALSE}
data[5,]
```
<br>

#### 2. What is the fifth lowest age?
##### I use unique() to sort out duplicate ages and arrange() to arrange the data based on ascending order
```{r,message=FALSE}
temp<-unique(arrange(data,age))
temp[5,]
```
<br>

#### 3. Extract the five lowest ages together
##### I use data[from:until,] to print out from a certain row number to a certain row number.
```{r,message=FALSE}
temp[1:5,]
```
<br>

#### 4. Get the five highest ages by first sorting them in decreasing order first.
##### Here I use sort and set the decreasing into TRUE so it returns. I also unique the data to remove the duplicate ages.
```{r,message=FALSE}
temp2<-sort(unique(data$age),decreasing = TRUE)
temp2[1:5]
```
<br>

#### 5. What is the average (mean) age?
##### I use mean() to get the average age.
```{r,message=FALSE}
mn<-mean(data$age)
mn
```
<br>

#### 6. What is the standard deviation of ages?
##### I use sd() that's already provided by r to find the standard deviation.
```{r,message=FALSE}
sd(data$age)
```
<br>


#### 7. Make a new variable called age_diff, with the difference between each age and the mean age
##### Here i use abs() so all the result will be positive since it would be make more sense to find the difference. To calculate the difference I just calculate it by age-average_age
```{r,message=FALSE}
data$age_diff<-abs(data$age-mn)
head(data,5)
```
<br>

#### 8. What is the average “difference between each age and the mean age”?
##### I use mean() to find the average
```{r,message=FALSE}
av<-mean(data$age_diff)
av
```
<br>

#### 9. Visualize the raw data as we did in class: (a) histogram, (b) density plot, (c) boxplot+stripchart
##### 
Histogram is used to summarize discrete or continuous data that are measured on an interval scale.
```{r,message=FALSE}
hist(data$age)
```
<br>

##### Density plots are used to observe the distribution of a variable in a dataset.
```{r,message=FALSE}
d<-density(data$age)
plot(d,main="Density of Age")
```
<br>

##### Box plots are used to show distributions and comparison between multiple groups. A strip chart can be used to visualize dozens of time series at once.
```{r,message=FALSE}
boxplot(data$age, horizontal = TRUE)
stripchart(data$age, method = "stack", add = TRUE)
```
<br>