---
title: "HW12"
author: '109006206'
output: pdf_document
---

## Set Working Directories & Reading Files
<br>

```{r,message=FALSE,out.width="80%"}
library(ggplot2)
setwd("/Users/olivia/Documents/Documents/Study/Semester 6/BACS/HW12")
cars<-read.table("auto-data.txt", header=FALSE, na.strings = "?")
names(cars) <- c("mpg", "cylinders", "displacement", "horsepower", "weight", 
                 "acceleration", "model_year", "origin", "car_name")
cars_log <- with(cars, data.frame(log(mpg), log(cylinders), log(displacement), log(horsepower), log(weight), log(acceleration), model_year, origin))

```

## QUESTION 1
<br>

#### A) Let’s visualize how weight might moderate the relationship between acceleration and mpg:

#### 1) Create two subsets of your data, one for light-weight cars (less than mean weight) and one for heavy cars (higher than the mean weight)

\
```{r,message=FALSE,out.width="80%"}
light_cars<-cars_log[cars_log$log.weight. < log(mean(cars$weight)), ]
heavy_cars<-cars_log[cars_log$log.weight. >= log(mean(cars$weight)), ]
```

#### 2) Create a single scatter plot of acceleration vs. mpg, with different colors and/or shapes for light versus heavy cars

\
```{r,message=FALSE,out.width="60%"}
ggplot() +
  geom_point(data = light_cars, aes(x = log.acceleration., y = log.mpg., color = "Light", shape = "Light")) +
  geom_point(data = heavy_cars, aes(x = log.acceleration., y = log.mpg., color = "Heavy", shape = "Heavy")) +
  scale_color_manual(values = c(Light = "red", Heavy = "blue")) +
  scale_shape_manual(values = c(16, 17)) +
  labs(x = "Acceleration", y = "MPG", color = "Weight", shape = "Weight")
```

#### 3) Draw two slopes of acceleration-vs-mpg over the scatter plot: one slope for light cars and one slope for heavy cars


\
```{r,message=FALSE,out.width="60%"}
ggplot() +
  geom_point(data = light_cars, aes(x = log.acceleration., y = log.mpg., color = "Light", shape = "Light")) +
  geom_point(data = heavy_cars, aes(x = log.acceleration., y = log.mpg., color = "Heavy", shape = "Heavy")) +
  geom_smooth(data = light_cars, aes(x = log.acceleration., y = log.mpg.), method = "lm", se = FALSE,fullrange=T, color = "red") +
  geom_smooth(data = heavy_cars, aes(x = log.acceleration., y =log.mpg.), method = "lm", se = FALSE,fullrange=T, color = "blue") +
  scale_color_manual(values = c(Light = "red", Heavy = "blue")) +
  scale_shape_manual(values = c(16, 17)) +
  labs(x = "Acceleration", y = "MPG", color = "Weight", shape = "Weight")
```

#### B) Report the full summaries of two separate regressions for light and heavy cars where log.mpg. is dependent on log.weight., log.acceleration., model_year and origin

\
```{r,message=FALSE,out.width="80%"}
light_regr <- lm(log.mpg. ~ log.weight. + log.acceleration. + model_year + origin, data = light_cars)
summary(light_regr)
heavy_regr <- lm(log.mpg. ~ log.weight. + log.acceleration. + model_year + origin, data = heavy_cars)
summary(heavy_regr)
```

#### C) Using your intuition only: What do you observe about light versus heavy cars so far?

**Answer** : Based on the scatter plot, both light and heavy cars have the same characteristic as the acceleration increases, the mpg also increases. 


\newpage

## QUESTION 2
<br>

#### A) Considering weight and acceleration, use your intuition and experience to state which of the two variables might be a moderating versus independent variable, in affecting mileage.

\
**Answer** : Acceleration is likely to be a moderating variable

#### B) Use various regression models to model the possible moderation on log.mpg.:

#### 1) Report a regression without any interaction terms

\
```{r,message=FALSE,out.width="80%"}
regr_mod <- lm(log.mpg. ~ log.weight. + log.acceleration. + model_year + factor(origin), data = cars_log)
summary(regr_mod)
```

#### 2) Report a regression with an interaction between weight and acceleration

\
```{r,message=FALSE,out.width="80%"}
regr_int <- lm(log.mpg. ~ log.weight. + log.acceleration. + model_year + factor(origin) + log.weight.*log.acceleration., data = cars_log)
summary(regr_int)
```

#### 3) Report a regression with a mean-centered interaction term

\
```{r,message=FALSE,out.width="80%"}
weight_mc <- scale(cars_log$log.weight., center=TRUE, scale=FALSE)
acceleration_mc <- scale(cars_log$log.acceleration., center=T, scale=F)
mpg_mc <- scale(cars_log$log.mpg., center=T, scale=F)
regr_interaction <- summary(lm(mpg_mc ~ weight_mc + acceleration_mc + weight_mc*acceleration_mc))
regr_interaction
```

#### 4) Report a regression with an orthogonalized interaction term

\
```{r,message=FALSE,out.width="80%"}
weight_acc <- cars_log$log.weight.*cars_log$log.acceleration.
interaction_regr <- lm(weight_acc ~ cars_log$log.weight. + cars_log$log.acceleration.)
interaction_ortho <- interaction_regr$residuals
summary(lm(log.mpg.~log.weight.+log.acceleration.+interaction_ortho,data=cars_log))
```

#### C) For each of the interaction term strategies above (raw, mean-centered, orthogonalized) what is the correlation between that interaction term and the two variables that you multiplied together?

1. Raw

\
```{r,message=FALSE,out.width="80%"}
cor(cars_log$log.weight., cars_log$log.weight. * cars_log$log.acceleration.)
cor(cars_log$log.acceleration., cars_log$log.weight. * cars_log$log.acceleration.)
```
\

2. Mean Centered

\
```{r,message=FALSE,out.width="80%"}
cor(weight_mc, weight_mc*acceleration_mc)
cor(acceleration_mc, weight_mc*acceleration_mc)
```
\

3. Orthogonalized

\
```{r,message=FALSE,out.width="80%"}
cor(interaction_ortho, cars_log$log.weight.)
cor(interaction_ortho, cars_log$log.acceleration.)
```

\newpage

## QUESTION 3
<br>

#### A) Let’s try computing the direct effects first:

#### 1) Model 1: Regress log.weight. over log.cylinders. only

\
```{r,message=FALSE,out.width="80%"}
model1 <- lm(log.weight. ~ log.cylinders., data = cars_log)
summary(model1)
```
\
**Answer** : Yes, based on the p-value it has a significant direct effect on weight


#### 2) Model 2: Regress log.mpg. over log.weight. and all control variables

\
```{r,message=FALSE,out.width="80%"}
model2 <- lm(log.mpg. ~ log.weight. + log.acceleration. + model_year + factor(origin), data = cars_log)
summary(model2)
```
\
**Answer** : Yes, based on the p-value weight has a significant direct effect on mpg

#### B) What is the indirect effect of cylinders on mpg?

\
```{r,message=FALSE,out.width="80%"}
indirect_effect <- model1$coefficients[2]*model2$coefficients[2]
indirect_effect
```

#### C) Let’s bootstrap for the confidence interval of the indirect effect of cylinders on mpg


#### 1) Bootstrap regression models 1 & 2, and compute the indirect effect each time: What is its 95% CI of the indirect effect of log.cylinders. on log.mpg.?

\
```{r,message=FALSE,out.width="80%"}
boot_mediation <- function(model1, model2, dataset) {
  boot_index <- sample(1:nrow(dataset), replace=TRUE)
  data_boot <- dataset[boot_index, ]
  regr1 <- lm(model1, data_boot)
  regr2 <- lm(model2, data_boot)
  return(regr1$coefficients[2] * regr2$coefficients[2])
}

set.seed(42)
indirect <- replicate(2000,boot_mediation(model1, model2, cars_log))
quantile(indirect, probs=c(0.025, 0.975))
```

\newpage

#### 2) Show a density plot of the distribution of the 95% CI of the indirect effect

\
```{r,message=FALSE,out.width="80%"}
plot(density(indirect))
abline(v=quantile(indirect, probs=c(0.025, 0.975)),lty=2)
```

