---
title: "HW3"
author: '109006206'
date: "helped by : 109060082"
output: pdf_document
header-includes:
  - \usepackage{setspace}
  - \onehalfspacing
  - \usepackage{fancyhdr}
  - \fancyhead[r]{109006206}
  - \pagestyle{fancy}
---

## Question 1
**a) Create a normal distribution (mean=940, sd=190) and standardize it (let’s call it rnorm_std)**
\
      **i) What should we expect the mean and standard deviation of rnorm_std to be, and why?**
      \
      \
      **Answer:**
      \
      We should expect the mean is 0 and standard deviation is 1 because it was standardized.
```{r,message=FALSE,out.width="80%"}
standardize <-function(numbers) {
  numbers <-(numbers -mean(numbers)) / sd(numbers)
  return(numbers)
  }
temp <- rnorm(1000,mean = 940, sd = 190)
rnorm_std <- standardize(temp)
print(paste("Mean :",mean(rnorm_std)))
print(paste("Standard Deviation:",sd(rnorm_std)))
```
\newpage
      **ii) What should the distribution (shape) of rnorm_std look like, and why?**
      \
      \
      **Answer:**
      \
      It should also be a normal distribution since "Standardization" shifts and scales a distribution and
      \
      won't change it's shape.
```{r,message=FALSE,out.heigth="50%"}
par(mfrow=c(1,2)) 
plot(density(temp),lwd=2,col="red",main="Before Standarized")
plot(density(rnorm_std),lwd=2,col="blue",main="After Standarized")
```
\
\
\
      **iii) What do we generally call distributions that are normal and standardized?**
      \
      \
      **Answer:**
      \
      It's called Standard Normal Distributions
      \
\
**b) Create a standardized version of minday discussed in question 3 (let’s call it minday_std)**
```{r,message=FALSE,out.heigth="50%"}
bookings <- read.table("first_bookings_datetime_sample.txt", header=TRUE)
hours  <- as.POSIXlt(bookings$datetime, format="%m/%d/%Y %H:%M")$hour
mins   <- as.POSIXlt(bookings$datetime, format="%m/%d/%Y %H:%M")$min
minday <- hours*60 + mins
```

```{r,message=FALSE,out.heigth="50%"}
minday_std<-standardize(minday)
```
\newpage
      **i) What should we expect the mean and standard deviation of minday_std to be, and**
      \
      **why?**
      \
      \
      **Answer:**
      \
      We should expect the mean is 0 and standard deviation is 1 because it was standardized
\
```{r,message=FALSE,out.width="80%"}
print(paste("Mean :",mean(minday_std)))
print(paste("Standard Deviation:",sd(minday_std)))
```
\
      **ii) What should the distribution of minday_std look like compared to minday, and why?**
```{r,message=FALSE,fig.dim = c(8, 4)}
par(mfrow=c(1,2)) 
plot(density(minday),lwd=3,col="red",main="Before Standardized")
abline(v=mean(minday),lwd = 2)
lines <- seq(-3,-1)
abline(v=mean(minday)-sd(minday)*lines,lty = "dashed")
lines2 <- seq(1,3)
abline(v=mean(minday)-sd(minday)*lines2,lty = "dashed")
plot(density(minday_std),lwd=3,col="blue",main="After Standardized")
abline(v=mean(minday_std),lwd = 2)
lines <- seq(-3,-1)
abline(v=mean(minday_std)-sd(minday_std)*lines,lty = "dashed")
lines2 <- seq(1,3)
abline(v=mean(minday_std)-sd(minday_std)*lines2,lty = "dashed")
```
\newpage

## Question 2
**a)Simulate 100 samples (each of size 100), from a normally distributed population of 10,000:**
\center
**plot_sample_ci(num_samples = 100, sample_size = 100, pop_size=10000, distr_func=rnorm, mean=20, sd=3)**
\raggedright
      **i) How many samples do we expect to NOT include the population mean in its 95% CI?**
      \raggedright
      **Answer:**

```{r,message=FALSE,out.heigth="50%"}
library("compstatslib")
n_samples<-100
cat("Number of samples we expect to NOT include the population mean in its 95% CI :"
    ,n_samples*0.05)
```
\
      **ii) How many samples do we expect to NOT include the population mean in their 99% CI?**
      \raggedright
      **Answer:**
      
```{r,message=FALSE,out.heigth="50%"}
n_samples<-100
cat("Number of samples we expect to NOT include the population mean in its 99% CI :"
    ,n_samples*0.01)
```

**b) Rerun the previous simulation with the same number of samples, but larger sample size (sample_size=300):**
\
      **i) Now that the size of each sample has increased, do we expect their 95% and 99% CI to**
      **become wider or narrower than before?**
```{r,message=FALSE,out.heigth="50%"}
plot_sample_ci(num_samples = 100, sample_size = 100, pop_size=10000, 
               distr_func=rnorm, mean=20, sd=3)
plot_sample_ci(num_samples = 100, sample_size = 300, pop_size=10000, 
               distr_func=rnorm, mean=20, sd=3)
```
We can see that the 95% and 99% CI become more narrower as the sample size increases
\newpage
      **ii) This time, how many samples (out of the 100) would we expect to NOT include**
      \
      **the population mean in its 95% CI?**
```{r,message=FALSE,out.heigth="30%"}
n_samples<-100
cat("Number of samples we expect to NOT include the population mean in its 95% CI :"
    ,n_samples*0.05)
```
\
**c) If we ran the above two examples (a and b) using a uniformly distributed population (specify parameter distr_func=runif for plot_sample_ci), how do you expect your answers to (a) and (b) to change, and why?**
\
```{r,message=FALSE,out.heigth="30%"}
plot_sample_ci(num_samples = 100, sample_size = 100, pop_size=10000, 
               distr_func=runif)
plot_sample_ci(num_samples = 100, sample_size = 300, pop_size=10000, 
               distr_func=runif)
```
\
\

## Question 3

\
**a) What is the “average” booking time for new members making their first restaurant booking?**
\
      **i) Use traditional statistical methods to estimate the population mean of minday, its**
      \
      **standard error, and the 95% confidence interval (CI) of the sampling means**
```{r,message=FALSE,out.heigth="30%"}
mean_minday<-mean(minday)
sd_error<-sd(minday)/(length(minday)^0.5)
ci95_low <- mean_minday-(1.96*sd_error)
ci95_high<- mean_minday+(1.96*sd_error)
cat("Mean :",mean_minday)
cat("Standard Deviation Error :",sd_error)
cat("95% Confidence Interval :",ci95_low," to ",ci95_high)
```
\newpage
      **ii) Bootstrap to produce 2000 new samples from the original sample**
```{r,message=FALSE,out.heigth="30%"}
resamples<-replicate(2000,sample(minday,length(minday),replace = TRUE))
```
\
      **iii) Visualize the means of the 2000 bootstrapped samples**
```{r,message=FALSE,out.heigth="30%"}
plot(density(minday), lwd=0, ylim=c(0, 0.009))
# Draws lines for each sampling mean
plot_resample_density<-function(sample_i) {
  lines(density(sample_i), col=rgb(0.0,1, 0.0, 0.01))
  return(mean(sample_i))
  }
sample_means<-apply(resamples, 2, FUN=plot_resample_density)
abline(v=mean(sample_means),col="blue",lwd=2)
abline(v=mean(minday),lty="dashed",lwd=3)
```
\
      **iv) Estimate the 95% CI of the bootstrapped means using the quantile function**
```{r,message=FALSE,out.heigth="30%"}
quantile(sample_means, probs=c(0.025, 0.975))
```
\newpage
**b) By what time of day, have half the new members of the day already arrived at their restaurant?**
\
      **i) Estimate the median of minday**
```{r,message=FALSE,out.heigth="30%"}
median(minday)
```
      **ii) Visualize the medians of the 2000 bootstrapped samples**
```{r,message=FALSE,out.heigth="30%"}
resamples2<-replicate(2000,sample(minday,length(minday),replace = TRUE))
plot(density(minday), lwd=0, ylim=c(0, 0.009))
# Draws lines for each sampling mean
plot_resample_density2<-function(sample_i) {
  lines(density(sample_i), col=rgb(0.0, 1, 0.0, 0.01))
  return(median(sample_i))
  }
sample_means2<-apply(resamples2, 2, FUN=plot_resample_density2)
abline(v=median(sample_means2), lwd=2)
abline(v=median(minday), lty="dashed")
```
      **iii) Estimate the 95% CI of the bootstrapped medians using the quantile function**
```{r,message=FALSE,out.heigth="30%"}
quantile(sample_means2, probs=c(0.025, 0.975))
```

