---
title: "HW7"
author: '109006206'
output: pdf_document
header-includes:
  - \usepackage{setspace}
  - \onehalfspacing
  - \usepackage{fancyhdr}
  - \fancyhead[r]{109006206}
  - \pagestyle{fancy}
---
## Set Working Directories & Reading Files
<br>
```{r,message=FALSE,out.width="80%"}
setwd("/Users/olivia/Documents/Documents/Study/Semester 6/BACS/HW7")
require("ggplot2")
require("tidyverse")
require("FSA")
Media_1 <- read.csv("pls-media1.csv")
Media_2 <- read.csv("pls-media2.csv")
Media_3 <- read.csv("pls-media3.csv")
Media_4 <- read.csv("pls-media4.csv")
```


## QUESTION 1
<br>

#### A) What are the means of viewers’ intentions to share (INTEND.0) on each of the four media types?
\
```{r,message=FALSE,out.width="80%"}
m1_mean <- mean(Media_1$INTEND.0)
m2_mean <- mean(Media_2$INTEND.0)
m3_mean <- mean(Media_3$INTEND.0) 
m4_mean <- mean(Media_4$INTEND.0)
```
m1_mean : `r mean(Media_1$INTEND.0)`
\
m2_mean : `r mean(Media_2$INTEND.0)`
\
m3_mean : `r mean(Media_3$INTEND.0)`
\
m4_mean : `r mean(Media_4$INTEND.0)`

#### B) Visualize the distribution and mean of intention to share, across all four media.
\
```{r,message=FALSE,out.width="80%"}
m <-rbind(Media_1,Media_2,Media_3,Media_4)
plot(density(Media_2$INTEND.0),col = 'pink',main="Distribution Across Four Media",xlab="Media",ylab="Intention.0")
lines(density(Media_1$INTEND.0),col = 'lightblue')
lines(density(Media_3$INTEND.0),col = 'lightgreen')
lines(density(Media_4$INTEND.0),col = 'lightgrey')
legend('topright',inset=0.001,c("Media 1","Media 2","Media 3","Media 4"),lty=1,col=c('lightblue','pink','lightgreen','lightgrey'),cex = 0.5)
```
\

#### C) From the visualization alone, do you feel that media type makes a difference on intention to share?
\
**Answer** : I think the media have some affect to the intention to share based on the visualization

\newpage

## QUESTION 2

#### A) State the null and alternative hypotheses when comparing INTEND.0 across four groups in ANOVA
\
Null Hypothesis: there is no difference in the means of INTEND.0 between the four groups 
\
(H0: $\mu$ 1 = $\mu$ 2 = $\mu$ 3 = $\mu$ 4)
\
Alternative Hypothesis : there is a significant difference in the means of INTEND.0 (Not all means are equal)

#### B) Let’s compute the F-statistic ourselves:

#### i) Show the code and results of computing MSTR, MSE, and F
\
```{r,message=FALSE,out.width="80%"}
mbind <- list(Media_1$INTEND.0,Media_2$INTEND.0,Media_3$INTEND.0,Media_4$INTEND.0)
k <- length(mbind)
n <- 4
grandmean<- mean(sapply(mbind,mean))
#MSTR
SSTR<-function(){
  (nrow(Media_1) * (mean(Media_1$INTEND.0) - grandmean)^2)+
  (nrow(Media_2) * (mean(Media_2$INTEND.0) - grandmean)^2)+
  (nrow(Media_3) * (mean(Media_3$INTEND.0) - grandmean)^2)+
  (nrow(Media_4) * (mean(Media_4$INTEND.0) - grandmean)^2)
}
df_mstr<-n-1
mstr<-SSTR()/df_mstr
mstr

#MSE
SSE<-function(){
    ((nrow(Media_1)-1) * (sd(Media_1$INTEND.0)^2))+
    ((nrow(Media_2)-1) * (sd(Media_2$INTEND.0)^2))+
    ((nrow(Media_3)-1) * (sd(Media_3$INTEND.0)^2))+
    ((nrow(Media_4)-1) * (sd(Media_4$INTEND.0)^2))
}
nt <- nrow(Media_1)+nrow(Media_2)+nrow(Media_3)+nrow(Media_4)
df_mse <- nt - k
mse <- SSE()/df_mse
mse

f_value <-mstr/mse
f_value
```
**MSTR** : `r mstr`
\
**MSE** : `r mse`
\
**F** : `r f_value`

#### ii) Compute the p-value of F, from the null F-distribution; is the F-value significant? If so, state your conclusion for the hypotheses.
\
```{r,message=FALSE,out.width="80%"}
p_value<-pf(f_value, df_mstr, df_mse, lower.tail=FALSE)
```
Since the p-value is `r p_value` so we can conclude that the F-value is not significant
\

#### C) Conduct the same one-way ANOVA using the aov() function in R – confirm that you got similar results.
\
```{r,message=FALSE,out.width="80%"}
m_new1 <- rbind(Media_1,Media_2,Media_3,Media_4)
m_new1 <- m_new1[1:2]
oneway.test(m_new1$INTEND.0 ~ factor(m_new1$media),var.equal = TRUE)
anova_model <- aov(m_new1$INTEND.0 ~ factor(m_new1$media))
summary(anova_model)
```


#### D) Regardless of your conclusions, conduct a post-hoc Tukey test (feel free to use the TukeyHSD() function included in base R) to see if any pairs of media have significantly different means – what do you find?
\
```{r,message=FALSE,out.width="80%"}
TukeyHSD(anova_model, conf.level = 0.05)
```
**Answer** : As we can see from the data we could say there is no different means at $\alpha$ = .05

#### E) Do you feel the classic requirements of one-way ANOVA were met?
\
1. The variance of the response variables is the same for all treatments/populations
\
**Answer** : As we can see from the data below we could conclude that there are some differences in the variance
```{r,message=FALSE,out.width="80%"}
Variance<- c(var(Media_1$INTEND.0),
var(Media_2$INTEND.0),
var(Media_3$INTEND.0),
var(Media_4$INTEND.0))
Variance
```
\
2. Each treatment/population’s response variable is normally distributed
```{r,message=FALSE,out.width="80%"}
shapiro.test(Media_1$INTEND.0)
shapiro.test(Media_2$INTEND.0)
shapiro.test(Media_3$INTEND.0)
shapiro.test(Media_4$INTEND.0)
```
\
**Answer** : As we can see from the result the second data p-value is bigger than 0.05 which indicates that the data is normally distributed.
\
\
**In conclusion, the classic requirements of one-way ANOVA were not met.**
\newpage

## QUESTION 3

#### A) State the null and alternative hypotheses
\
Null Hypothesis : All media would give similar value if randomly drawn from them 
\
(H0: $\mu$ 1 = $\mu$ 2 = $\mu$ 3 = $\mu$ 4)
\
Alternative Hypothesis : At least one media would give a larger value than another if randomly drawn

#### B) Let’s compute (an approximate) Kruskal Wallis H ourselves (use the formula we saw in class or another formula might have found at a reputable website/book):

#### i) Show the code and results of computing H
\
```{r,message=FALSE,out.width="80%"}
mkw <- m[1:2]
ranks <- rank(mkw$INTEND.0)
N<-length(ranks)
group_ranks <- split(ranks,f = mkw$media)

sum_ranks <-sapply(group_ranks,sum)
length_ranks<- (sapply(group_ranks,length))
Total<- sum((sum_ranks^2)/length_ranks)

H<- (((12/(N*(N+1))) * Total) - 3*(N+1))
H
```
**H** : `r H`

#### ii) Compute the p-value of H, from the null chi-square distribution; is the H value significant? If so, state your conclusion of the hypotheses.
\
```{r,message=FALSE,out.width="80%"}
kw_p<-1 -pchisq(H, df=k-1)
kw_p
```
**P-Value of H** : `r kw_p`
\
**Answer** : In conclusion, we should reject the null hypothesis since the p-value < 0.05
\newpage

#### C) Conduct the same test using the kruskal.wallis() function in R – confirm that you got similar results.
\
```{r,message=FALSE,out.width="80%"}
kruskal.test(mkw$INTEND.0 ~ mkw$media, mkw)
```

#### D) Regardless of your conclusions, conduct a post-hoc Dunn test to see if the values of any pairs of media are significantly different – what are your conclusions?
\
```{r,message=FALSE,out.width="80%"}
dunnTest(INTEND.0~media, data = mkw , method = "bonferroni")
```
\
**Answer** : At $\alpha$ = .05, we can conclude that media 2 and media 4 are statistically significantly different from each other (P.adj = 0.0474) 
